---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# nansen

[![CRAN status](https://www.r-pkg.org/badges/version/galalh)](https://cran.r-project.org/package=galalh)

`nansen` is an R package designed to facilitate access to the United Nations High Commissioner for Refugees (UNHCR) Refugee Data Finder. It provides an easy-to-use interface to the database, which covers forcibly displaced populations, including refugees, asylum-seekers, internally displaced people, stateless people, and others over a span of more than 70 years.

This package provides data from three major sources:

1. Data from UNHCR's annual statistical activities dating back to 1951.
2. Data from the United Nations Relief and Works Agency for Palestine Refugees in the Near East (UNRWA), specifically for registered Palestine refugees under UNRWA’s mandate.
3. Data from the Internal Displacement Monitoring Centre (IDMC) on people displaced within their country due to conflict or violence.

## Datasets

The `nansen` package includes eight datasets:

1. `population`: Data on forcibly displaced and stateless persons by year, including refugees, asylum-seekers, returned refugees, internally displaced persons (IDPs), and stateless persons.

2. `idmc`: Data from the Internal Displacement Monitoring Centre on the total number of IDPs due to conflict and violence.

3. `asylum_applications`: Data on asylum applications including procedure type and application type.

4. `asylum_decisions`: Data on asylum decisions, including recognised, rejected, and otherwise closed claims.

5. `demographics`: Demographic and sub-national data, where available, including age and sex disaggregation.

6. `solutions`: Data on durable solutions for refugees and IDPs.

7. `unrwa`: Data on registered Palestine refugees under UNRWA’s mandate.

8. `countries`: Country codes, names, UN major areas, and UNHCR regional bureaux/operations.

Please check the individual dataset description for more details on the fields each dataset provides.

## Installation

Install either from CRAN with:

```{r, eval = FALSE}
install.packages("nansen")
```

Or retrieve the development version from Github with:

```{r, eval = FALSE}
remotes::install_github("galalH/nansen")
```

## Usage

Here are some examples of how you can use the `nansen` package.

```{r}
# Load the package
library(nansen)
str(nansen::population)

# get the total number of refugees (including refugee-like) from Sudan in Chad in 2021
nansen::population |>
  subset(year == 2021 & coa_iso == "TCD" & coo_iso == "SDN",
         select = refugees)
```

## Why `nansen`?

The package is named as homage to [Fridtjof Wedel-Jarlsberg Nansen](https://en.wikipedia.org/wiki/Fridtjof_Nansen), 
a Norwegian explorer, scientist, diplomat, and humanitarian who significantly contributed to assisting displaced persons following World War I. Serving as the League of Nations High Commissioner for Refugees, Nansen developed the "Nansen Passport", a critical tool that allowed stateless persons to cross borders legally. His groundbreaking work earned him the Nobel Peace Prize in 1922 and laid the foundation for modern refugee law.

By naming this package `nansen`, we pay tribute to his lasting legacy in service of forcibly displaced people worldwide. Our hope is that this package aids in continuing this legacy of understanding, assisting, and finding solutions for displaced populations worldwide.

## License

This package is released under the [CC0 License](https://creativecommons.org/share-your-work/public-domain/cc0/)
